---
title: "Dispersal Simulation using RangeShiftR"
output:
  pdf_document: default
  html_document: default
  word_document: default
---
```{r, echo=TRUE, warning=FALSE, message=FALSE}
library(RangeShiftR)
require(sp)
require(rgdal)
require(sfheaders)
require(raster)
require(SDMSelect)
require(tidyverse)
```


# Introduction

The goal of the analysis is to simulate dispersal of _Cakile edentula_ following introductions at the ports of "Sydney", "Melbourne", "Brisbane" and "Perth".

Steps taken:

1. Using the [BCCVL](https://bccvl.org.au/) platform, and "occurrence data" for _Cakile edentula_ in its native habitat (in USA) and the climate data "CRUclim (global), current climate (1976-2005), 30 arcmin (~50 km)" as input, predicted a species distribution map for the suitability of _Cakile edentula_.  The output is a CSV file with the suitability probabilities (0 - 1) for different points.

2. Transformed the data on suitability from BCCVL into a landscape map file *land.txt*.

```{r, echo=TRUE, warning=FALSE, message=FALSE}
#### suitability map

l = raster("Inputs/proj_current_cakile.edentula_glm_unconstrained.tif")
l2 = raster::resample(l, raster(ext=extent(extent(l)), resolution=2,crs=projection(l)))

raster::writeRaster(l2, format="ascii", filename = "Inputs/climate_suitabilitya", NAflag = -9, overwrite = T, bylayer = T, datatype = "FLT4S")


lp = gplot_data(raster("Inputs/climate_suitabilitya.asc")) %>% filter(!is.na(value))
```


3. Identified the map locations for the ports of "Sydney", "Melbourne", "Brisbane" and "Perth" and randomly assigned then a number [$norm (\mu = 25, \sigma = 5)$] of _Cakile edentula_ introductions. Transformed this data into a species distribution map file *introduction.txt*

```{r, echo=TRUE, warning=FALSE, message=FALSE}
introduction = read.csv("Inputs/aus_ports3.csv")[, 1:3]
introduction$Individuals = as.numeric(introduction$Individuals)

a3 = introduction %>% filter(!is.na(Individuals)) %>% mutate("locations" =  c("Syd", "Mel", "Mel", "Bris", "Perth")) %>% na.omit()

coordinates(introduction) = ~ lon + lat
proj4string(introduction)=proj4string(l) # set it to lat-long
introduction = spTransform(introduction,proj4string(l))

e <- extent(l)
r <- raster(e, ncols = 720, nrows = 360)

# you need to provide a function 'fun' for when there are multiple points per cell
introduction2 <- rasterize(introduction, r, fun = mean)

introduction3 = raster::resample(introduction2[[2]], raster(ext=extent(extent(introduction2)), resolution=2,crs=projection(introduction)))


raster::writeRaster(introduction3, format="ascii", filename = "Inputs/news_locations.asc", NAflag = -9, overwrite = T, datatype = "INT2U")

ggplot( ) + 
  geom_tile(data = lp , aes(x = x, y = y, fill = value)) + 
  scale_fill_gradient("suitability", low = "white", high = "black") +
  theme_grey() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + 
  geom_point(data = a3, aes(x = lon, y = lat, color = locations))
```

The chart above displays the climate suitability for _Cakile edentula_ in the selected region that overlaps between the native area in the USA and Australia. From the map, we observe that suitability is only relatively high in the native region and very low everywhere else. The introduction points for this simulation are also shown and we note that they are all regions with very low suitability.


4. Define the *landscape parameter* in RangeshiftR for simulation. We define the resolution at $1km$. We set *HabPercent = TRUE* since the data in our suitability map is a probability value between 0 and 1. We set the *K_or_DensDep = 2* to infer that a location with a probability of 1 would have a carrying capacity of 2. The carrying capacities for all locations are calculated based on the probability value and the 100 threshold. 

```{r, echo=TRUE, warning=FALSE, message=FALSE}
landscapes = ImportedLandscape(LandscapeFile = "climate_suitabilitya.asc", 
                          Resolution = 2, 
                          HabPercent = TRUE, 
                          K_or_DensDep = 100, 
                          SpDistFile = "news_locations.asc", 
                          SpDistResolution = 2)

```

5. We define the Species parameters which include:

- *demography*: I have made the assumption that the population dynamics are described by a female-only model with non-overlapping generations and defined the *maximum growth rate to 1.4*

```{r, echo=TRUE, warning=FALSE, message=FALSE}
demos <- Demography(Rmax = 1000, bc = 0.1, ReproductionType = 0)
demos
```

- *dispersal*: This contains information on *Emigration* probability, *Transport* modeled with a dispersal kernel and *Settlement* whose default options assume that an individual will die if it arrives in an unsuitable cell and settle if it’s suitable. 

We make the assumption that the *Emigration probability = 0.1*

We define *short range* and *long range transfer* at $20,000m$ with $probability = 0.5$ and $100,000m$ with $probability = 0.1$ respectively. This suggests that we have a 3rd possible dispersal distance equal to $0$ which has a probability of $1 - (0.5 + 0.1) = 0.4$. However, RangeShiftR  provides for either a single or double kernel to define distance. To fit this, I considered the odds between the short dispersal distance (20km) and the long distance (100km) and used this to recalculate their probabilities assuming the 3rd option of $0Km$ is not present.

$$P_{20km} = P_{20km}/{(P_{20km} + P_{100Km})} = 0.83$$

$$P_{100km} = P_{100km}/{(P_{20km} + P_{100Km})}=0.17$$






```{r, echo=TRUE, warning=FALSE, message=FALSE}
dist = matrix(c(20000, 100000, 0.83 ), ncol = 3, byrow = T)

disps <-  Dispersal(Emigration = Emigration(EmigProb = 0.9), Transfer =  DispersalKernel(DoubleKernel = T, Distances = dist)) + Settlement()
disps
```

6. Initialize the parameters for simulation

```{r, echo=TRUE, warning=FALSE, message=FALSE}
inits <- Initialise(InitType = 0, # = initialisation from a loaded species distribution map
                   SpType = 0,   # = all suitable cells within all distribution presence cells
                   InitDens = 2,
                   IndsHaCell = 1000) # = at carrying capacity

inits
```

7. Simulation parameters

We will run $1\ experiments$ with $10\ replicates$ simulating the spread of _Cakile edentula_ over $1000\ years$ and record the *population*, *occurrence* and *range* every $5\ years$

```{r, echo=TRUE, warning=FALSE, message=FALSE}
sims <- Simulation(Simulation = 1,
                  Years = 1000,
                  Replicates = 2,
                  OutIntRange = 1,
                  OutIntPop = 1,
                  OutIntOcc = 1)
sims
```

8. Combine all the predefined parameters into a master object and run the simulation. We will run two separate simulations, one for the short range dispersal and the other for the long range dispersal.

```{r, echo=TRUE, warning=FALSE, message=FALSE}
simulate <- RSsim() + landscapes + demos + disps + sims + inits
validateRSparams(simulate) ## check parameter validity
```

9. Run simulations

```{r, echo=TRUE, message=FALSE}
RunRS(RSparams = simulate, dirpath = paste0(getwd(), "/"))
```