
Simulation data
```{r}
db <- "postgres" 
host_db <- "postgres.cl0erzvvnfux.ap-northeast-1.rds.amazonaws.com"   
db_port <- 5432  
db_user <- "philipp"  
db_password <- '****' # enter the password here

con <- dbConnect(RPostgres::Postgres(), dbname = db, host=host_db, port=db_port, user=db_user, password=db_password) 

Simulation = dbReadTable(con, "simulated_population_lag_table_4_locations") |> dplyr::select(-row.names)
```

# proportion of Lag species

```{r, message=FALSE}
library(tidyverse)
(Simulation |> select(X..Lag.Species))$X..Lag.Species |> mean(na.rm = T)
```

Australia lag data

```{r}
AUS_data = dbReadTable(conn = con, "AUS_lag_results") |>
  mutate(lengthlag0 = as.numeric(lengthlag0), 
         lengthlag = as.numeric(lengthlag))
```


# proportion of Lag species

```{r}
(sum(AUS_data$increase0 == TRUE)/nrow(AUS_data) * 100) |> round() # with the 1st set of lag data
```


# mean lag lengths

```{r}
(AUS_data |> filter(increase0 == TRUE))$lengthlag0 |> mean(na.rm = T) # with the 1st set of lag data
```




# AUS vs Simulation

```{r}
t.test((AUS_data |> filter(increase0 == TRUE))$lengthlag0, Simulation$Mean.Lag.End) # with the 1st set of lag data
```


NZL lag data

```{r}
NZL_data = dbReadTable(conn = con, "NZL_lag_results") 
(sum(NZL_data$increase0 == TRUE, na.rm = T)/nrow(NZL_data) * 100) |> round() # with the 1st set of lag data
```


# mean lag lengths

```{r}
(NZL_data |> filter(increase0 == TRUE))$lengthlag0 |> mean() # with the 1st set of lag data
```



# NZL vs Simulation

```{r}
t.test((NZL_data|> filter(increase0 == TRUE))$lengthlag0, Simulation$Mean.Lag.End) # with the 1st set of lag data
```



lag lengths

```{r}
d1 = data.frame(lags = c(Simulation$Mean.Lag.End, AUS_data$lengthlag0, NZL_data$lengthlag0), 
                data = c(rep("Simulated", length(Simulation$Mean.Lag.End)), 
                         rep("Herbarium (AUS)", length(AUS_data$lengthlag0)),
                         rep("Herbarium (NZL)", length(NZL_data$lengthlag0)))) |> 
  na.omit()

# lag lengths
ggplot(d1, aes(x = lags, fill = data)) + 
  geom_histogram(aes(color = data, fill = data), position = "dodge", alpha = 0, bins = 20) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  ylab("Number of Lag Species") + xlab("Lag lengths")

```


```{r}
# AUS

# lagged species
AUS_data2 = AUS_data[AUS_data$increase0 == TRUE,]

# Getting range data for AUS species with lags
data3 = readr::read_csv("~/Desktop/Documents/GitHub/dispersal-simulation-changed-files/herbaria pop data/AUS-data-lag.csv")
data3$Species = tolower(data3$Species)

# from species identified to have lag, extract the residence time by aggregating the range data
data4 = data3[data3$Species %in% AUS_data2$species, ] |>
  group_by(Species) |> 
  summarise(residence = max(Year) - min(Year))

# merge the data with residence time to the lag data and compute proportion of lag length to residence time
AUS_data2 = merge(AUS_data2, data4, by.x = "species", by.y = "Species") |>
  mutate(`lag % of residence` = round(100 * lengthlag0/residence, 2)) 



# NZL
# lagged species
NZL_data2 = NZL_data[NZL_data$increase0 == TRUE,]
NZL_data2$species = tolower(NZL_data2$species)

# Getting range data for NZL species with lags
data5 = readr::read_csv("~/Desktop/Documents/GitHub/dispersal-simulation-changed-files/herbaria pop data/NZ-data-lag.csv")
data5$Species = tolower(data5$Species)

# from species identified to have lag, extract the residence time by aggregating the range data
data6 = data5[data5$Species %in% NZL_data2$species, ] |>
  group_by(Species) |> 
  summarise(residence = max(Year) - min(Year))

# merge the data with residence time to the lag data and compute proportion of lag length to residence time
NZL_data2 = merge(NZL_data2, data6, by.x = "species", by.y = "Species") |>
  mutate(`lag % of residence` = round(100 * lengthlag0/residence, 2)) 



d2 = data.frame(lags = c(Simulation$X..Lag.Length.of.Years.Assessed, AUS_data2$`lag % of residence`, NZL_data2$`lag % of residence`), 
                data = c(rep("Simulated", length(Simulation$X..Lag.Length.of.Years.Assessed)), 
                         rep("Herbarium (AUS)", length(AUS_data2$`lag % of residence`)),
                         rep("Herbarium (NZL)", length(NZL_data2$`lag % of residence`)))) |> 
  na.omit()
```


```{r}
# lag lengths
ggplot(d2, aes(x = lags, fill = data)) + 
  geom_histogram(aes(y = after_stat(c(
      count[group == 1]/sum(count[group == 1]),
      count[group == 2]/sum(count[group == 2]),
      count[group == 3]/sum(count[group == 3])
    ) * 100), color = data, color = data), 
    position = "dodge", alpha = 0, bins = 20) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  ylab("Number of Lag Species (Percentage)") + xlab("Lag lengths vs Residence time")
```

Correlation between number of lag events and the average duration of lag

```{r}
cor.test(Simulation$LagSpecies, Simulation$Mean.Lag.End)
```


95% confidence interval for the mean lag lengths in viirtual species

```{r}
mean = Simulation$Mean.Lag.End |> mean()
sd = Simulation$Mean.Lag.End |> sd()
t = qnorm(0.025, lower.tail = F)

lower_lim = mean - (t * sd) 
upper_lim = mean + (t * sd) 

# NZL neophytes within simulation confidence interval
(NZL_data2$lengthlag0[NZL_data2$lengthlag0 < round(upper_lim)] |> length()) / (NZL_data2$lengthlag0 |> length()) * 100
data5[data5$Species %in% NZL_data2$species, "Year"] |> min() # min year
data5[data5$Species %in% NZL_data2$species, "Year"] |> max() # max year


# AUS neophytes within simulation confidence interval
(AUS_data2$lengthlag0[AUS_data2$lengthlag0 < round(upper_lim)] |> length()) / (AUS_data2$lengthlag0 |> length()) * 100
data3[data3$Species %in% AUS_data2$species, "Year"] |> min() # min year
data3[data3$Species %in% AUS_data2$species, "Year"] |> max() # max year
```
###
_________________
Demographic parameters: "competition coefficient: _increase lags, _increase % lags :: _reduce lag length, _reduce % lag length"



Dispersal parameters: "emigration probability : _increase lags, _increase % lags :: _reduce % lag length"
                      "short-range probability: _increase % lags :: _reduce % lag length" 
                      "short-range dispersal: _reduce lags, _reduce % lags :: _increase lag length", _increase % lag length
                      
                       
                      

Landscape parameters: "carrying capacity: _reduce lags,  _reduce % lags"


Initialization and simulation parameters:  "relative saturation at the introduction location: _reduce lags :: _increase lag length",  _increase % lag length
                                            "residence time: _reduce lags :: _reduce % lag length"
                                            "replicates:  _reduce lags ::  _increase % lag length"
