*Summary of the parameters and lag values*

```{r}
# database table to store simulation output
db <- "postgres" 
host_db <- "postgres.cl0erzvvnfux.ap-northeast-1.rds.amazonaws.com"   
db_port <- 5432  
db_user <- "philipp"  
db_password <- '****' # enter the password here

con <- dbConnect(RPostgres::Postgres(), dbname = db, host=host_db, port=db_port, user=db_user, password=db_password) 

data = dbReadTable(con, "simulated_population_lag_table_4_locations")
# or local file: read.csv("lag results - Oct 6th 2022.csv")
View(data)
```


```{r}
library(tidyverse)
library(psych)
library(car)


# data = read_csv("lag_table.csv") |> mutate(short_range_distance = short_range_distance/1000, long_range_distance = long_range_distance/1000)
#data = read_csv("new_lag_results - new_lag_results.csv") |> mutate(short_range_distance = short_range_distance/1000, long_range_distance = long_range_distance/1000)


# introduction density
data = data |> mutate(density = no_introduced_per_cell / carrying_capacity) |> 
  rename(`Mean Lag End` = Mean.Lag.End, 
         `% Lag species` = X..Lag.Species,
         `% Lag Length of Years Assessed` = X..Lag.Length.of.Years.Assessed)

mean_suitability = aggregate(intro_suitability ~ intro_location, data = data, FUN = mean)
mean_suitability |>
  knitr::kable()


describe(data, fast = TRUE) |> round(2)  |> knitr::kable()
```


```{r}
hist(data$LagSpecies)
```





*lag frequency*

Regression coefficients

```{r}
data2 = data |>
  select(-c(intro_location, 
            `Mean Lag End`, `% Lag species`, 
            `% Lag Length of Years Assessed`))

`fit lag` = lm(LagSpecies ~ 
                 carrying_capacity +
                 competition_coefficient +
                 short_range_distance + 
                 emigration_probability + 
                 Year + Species + density, data2)
summary(`fit lag`)$coefficients |> knitr::kable()


```

Q-Q plot

```{r}
plot(residuals(`fit lag`), ylab = "Residuals")
abline(h = 0)
```

Added variable plot

```{r}
library(car)
avPlots(`fit lag`) 
```

# % lag species

Regression coefficients

```{r}
data2 = data |>
  select(-c(intro_location, 
            `Mean Lag End`, LagSpecies, 
            `% Lag Length of Years Assessed`))

`fit % lag` = lm(`% Lag species` ~ carrying_capacity + competition_coefficient + short_range_distance + short_range_probability + emigration_probability  + density , data2)
summary(`fit % lag`)$coefficients |> knitr::kable()


```

Q-Q plot

```{r}
plot(residuals(`fit % lag`), ylab = "Residuals")
abline(h = 0)
```

Added variable plot

```{r}
avPlots(`fit % lag`) 
```

# lag lengths

Regression coefficients

```{r}
data2 = data |>
  select(-c(intro_location, 
            `% Lag species`, LagSpecies, 
            `% Lag Length of Years Assessed`))

`fit year` = lm(`Mean Lag End` ~ competition_coefficient + short_range_distance + short_range_probability + emigration_probability + density, data2)
summary(`fit year`)$coefficients |> knitr::kable()
```

Q-Q plot

```{r}
plot(residuals(`fit year`), ylab = "Residuals")
abline(h = 0)
```

Added variable plot

```{r}
avPlots(`fit year`) 
```

Parameters with Lag lengths similar to the observations from herbaria data

```{r}
SE = summary(`fit year`)$coefficients[1, 2]
average = summary(`fit year`)$coefficients[1,1]
T_ = floor(summary(`fit year`)$coefficients[1,3])

confidence_interval = function(average, SE, T_){
  lower = (T_ * SE) + average
  upper = (-T_ * SE) + average
  list(lower, upper)
}

conf.values = confidence_interval(average = average, SE = SE, T_ = T_)

data2 = data |> mutate(short_range_distance = short_range_distance/1000, long_range_distance = long_range_distance/1000)

data2 = data2 |>  
  dplyr::select(-intro_location) |>
  dplyr::filter(Mean.Lag.End > 3)  |> 
  dplyr::filter (Mean.Lag.End < 92) 
```

Summary of Parameters with Lag lengths similar to the observations from herbaria data

```{r}
data2 |> filter(`Mean Lag End` > conf.values[[1]])  |> filter (`Mean Lag End` < conf.values[[2]]) |> dplyr::select(short_range_distance, emigration_probability, Year) |> describe(fast = T) |> knitr::kable()
```

# % lag years

Regression coefficients

```{r}
data2 = data |>
  select(-c(intro_location, 
            `% Lag species`, LagSpecies, 
            `Mean Lag End`))

`fit % year` = lm(`% Lag Length of Years Assessed` ~ competition_coefficient + short_range_distance + short_range_probability + emigration_probability + Year + Species + density, data2)

summary(`fit % year`)$coefficients |> knitr::kable()
```

Q-Q plot

```{r}
plot(residuals(`fit % year`), ylab = "Residuals")
abline(h = 0)
```

Added variable plot

```{r}
avPlots(`fit % year`) 
```


_____________
To calculate the distance between points:

```{r}
# area between coord points in square kilo metres
distance_m = function(data, lon, lat) {
  radius = 6371 # km (radius at equator)
  r = 2 * pi * radius / 360 # distance that corresponds to 1 degree at Equator
  
  lat_change = data$lat - lat
  lon_change = (data$lon - lon) * cos((data$lat + lat) * 0.5 * pi/180)  
  d = r * sqrt(lat_change * lat_change + lon_change * lon_change)
  
  return(d)
}

data <- read.csv("Inputs/Unknown.species_AllData_Full.csv")

# Sydney 1
ss = data |> filter(lon > 151 & lon < 153 & lat > -34 & lat < -33)
sydney_distance = distance_m(ss, 152.0886, -33.60759)
sydney_distance

# melbourne
mm = data |> filter(lon > 145 & lon < 146 & lat > -39 & lat < -38) 
melbourne_distance = distance_m(mm, 145.2532, -38.16456)
melbourne_distance

# Brisbane 1
bb = data |> filter(lon > 153 & lon < 155 & lat > -29 & lat < -27) 
brisbane_distance = distance_m(bb, 154.3671, -29.05063)
brisbane_distance

# Perth 1
pp = data |> filter(lon > 115 & lon < 116 & lat > -31 & lat < -30) 
perth_distance = distance_m(pp, 115.6329, -31.32911)
perth_distance
```

```{r}
bb$Location = "Brisbane"
pp$Location = "Perth"
ss$Location = "Sydney"
mm$Location = "Melbourne"

h = rbind(bb, pp, ss, mm) |> rename(suitability = Unknown.species_AllData_Full.logistic.values)
h2 = h |> select(Location, lon, lat, suitability) |> mutate(suitability = round(100 * suitability, 2))
distances = c(brisbane_distance, perth_distance, sydney_distance, melbourne_distance)

h2$`distance from port` = distances
h2 |> knitr::kable()

hd = aov(suitability ~ Location, h2)
summary(hd)
TukeyHSD(hd)$Location |> knitr::kable()
```
