---
title: "Cakile edentula Range Dispersal"
author: "Kwiz Research Services"
date: '2022-05-10'
output: word_document
---


```{r, echo=TRUE, warning=FALSE, message=FALSE}
library(RangeShiftR)
require(sp)
require(rgdal)
require(sfheaders)
require(raster)
require(SDMSelect)
require(tidyverse)
```


# Introduction

The goal of the analysis is to simulate dispersal of cakile edentula following introductions at the ports of "Sydney", "Melbourne", "Brisbane" and "Perth".

Steps taken:

1. Using the BCCVL platform, and "occurance data" for cakile edentula in its native and observed habitats (globally) and the climate data "CRUclim (global), current climate (1976-2005), 30 arcmin (~50 km)" as input, predicted a species distribution map for the suitability of cakile edentula.  The output is a CSV file with the suitability probabilities (0 - 1) for different points.

2. Transformed the data on suitability from BCCVL into a landscape map file *climate_suitabilitya*. Extract the probabilities with Australia only i.e

     extent(xmin = 110, xmax = 155, ymin = -45, ymax = -9)

```{r, echo=TRUE, warning=FALSE, message=FALSE}
#### suitability map from csv
AllData <- read.csv("Inputs/Unknown.species_AllData_Full.csv")
colnames(AllData)[1:2] = c("lon", "lat")
coordinates(AllData) = ~ lon + lat
proj4string(AllData)=CRS("+init=EPSG:4326") # set it to lat-long
AllData = spTransform(AllData,CRS("+init=EPSG:4326"))
e <- extent(110, 155, -45, -9)
r <- raster(e, ncols = 10000, nrows = 10000)
# you need to provide a function 'fun' for when there are multiple points per cell
AllData_a <- rasterize(AllData, r, fun = mean)
AllData2a = raster::resample(AllData_a[[2]], raster(ext= e, resolution = 0.2,crs=projection(AllData)))
raster::writeRaster(AllData2a * 100, format="ascii", filename = "Inputs/climate_suitability", NAflag = -9, overwrite=TRUE, bylayer = T, datatype = "INT4U")
```


3. We simulate a species distribution map such that we have 100 individuals each in 4 locations i.e the ports of "Sydney", "Melbourne", "Brisbane" and "Perth". Save this data into a species distribution map file *news_locations.asc*

```{r, echo=TRUE, warning=FALSE, message=FALSE}
introduction = read.csv("Inputs/aus_ports3.tsv", sep = "\t")[, 3:5] |> rename(lon = X, lat = Y, Individuals = NInds) |> distinct()
introduction$Individuals = as.numeric(introduction$Individuals)
a3 = introduction %>% filter(!is.na(Individuals)) %>% mutate("locations" =  c("Syd", "Mel", "Bris", "Perth")) %>% na.omit()
```

To calculate the distance between points:

```{r}
# area between coord points in square kilo metres
distance_m = function(data, lon, lat) {
  radius = 6371 # km (radius at equator)
  r = 2 * pi * radius / 360 # distance that corresponds to 1 degree at Equator
  
  lat_change = data$lat - lat
  lon_change = (data$lon - lon) * cos((data$lat + lat) * 0.5 * pi/180)  
  d = r * sqrt(lat_change * lat_change + lon_change * lon_change)
  
  return(d)
}
```



```{r, echo=TRUE, warning=FALSE, message=FALSE}
data <- read.csv("Inputs/Unknown.species_AllData_Full.csv")

# Sydney 1
coords = data |>dplyr::select(lon, lat)
coords$distances = distance_m(coords, 151.1989, -33.85151)
sydney = coords |> filter(distances == min(distances))

ss = data |> filter(lon == sydney$lon & lat == sydney$lat)
sydney_distance = sydney$distances
sydney_distance

ss1 = ss[1, ]
sydney = data.frame(lon = ss1$lon, lat = ss1$lat, Individuals = 1, locations = "Syd")

coordinates(sydney) = ~ lon + lat
proj4string(sydney)=proj4string(AllData) # set it to lat-long
sydney = spTransform(sydney,proj4string(AllData))
e <- extent(110, 155, -45, -9)
r <- raster(e, ncols = 10000, nrows = 10000)
# you need to provide a function 'fun' for when there are multiple points per cell
sydney2 <- rasterize(sydney, r, fun = mean)
sydney3 = raster::resample(sydney2[[2]], raster(ext= e, resolution= 1,crs=projection(sydney)))
raster::writeRaster(sydney3, format="ascii", filename = "Inputs/intro_sydney_port.asc", NAflag = -9, overwrite = T, datatype = "UTF-8")
```

```{r, echo=TRUE, warning=FALSE, message=FALSE}
# Melbourne 1
coords = data |>dplyr::select(lon, lat)
coords$distances = distance_m(coords, 144.937, -37.83878)
melbourne = coords |> filter(distances == min(distances))

mm = data |> filter(lon == melbourne$lon & lat == melbourne$lat)
melbourne_distance = melbourne$distances
melbourne_distance

  
mm1 = mm[1,]
Melbourne = data.frame(lon = mm1$lon, lat = mm1$lat, Individuals = 1, locations = "Mel")


coordinates(Melbourne) = ~ lon + lat
proj4string(Melbourne)=proj4string(AllData) # set it to lat-long
Melbourne = spTransform(Melbourne,proj4string(AllData))
e <- extent(110, 155, -45, -9)
r <- raster(e, ncols = 10000, nrows = 10000)
# you need to provide a function 'fun' for when there are multiple points per cell
Melbourne2 <- rasterize(Melbourne, r, fun = mean)
Melbourne3 = raster::resample(Melbourne2[[2]], raster(ext= e, resolution= 1,crs=projection(Melbourne)))
raster::writeRaster(Melbourne3, format="ascii", filename = "Inputs/intro_Melbourne_port.asc", NAflag = -9, overwrite = T, datatype = "UTF-8")
```

```{r, echo=TRUE, warning=FALSE, message=FALSE}
# Brisbane 1
coords = data |>dplyr::select(lon, lat)
coords$distances = distance_m(coords, 153.0956, -27.41807)
brisbane = coords |> filter(distances == min(distances))

bb = data |> filter(lon == brisbane$lon & lat == brisbane$lat)
brisbane_distance = brisbane$distances
brisbane_distance


bb1 = bb[1, ]
Brisbane = data.frame(lon = bb1$lon, lat = bb1$lat, Individuals = 1, locations = "Bris")

coordinates(Brisbane) = ~ lon + lat
proj4string(Brisbane)=proj4string(AllData) # set it to lat-long
Brisbane = spTransform(Brisbane,proj4string(AllData))
e <- extent(110, 155, -45, -9)
r <- raster(e, ncols = 10000, nrows = 10000)
# you need to provide a function 'fun' for when there are multiple points per cell
Brisbane2 <- rasterize(Brisbane, r, fun = mean)
Brisbane3 = raster::resample(Brisbane2[[2]], raster(ext= e, resolution= 1,crs=projection(Brisbane)))
raster::writeRaster(Brisbane3, format="ascii", filename = "Inputs/intro_Brisbane_port.asc", NAflag = -9, overwrite = T, datatype = "UTF-8")

```



```{r, echo=TRUE, warning=FALSE, message=FALSE}
# Perth 1
coords = data |>dplyr::select(lon, lat)
coords$distances = distance_m(coords, 115.8834, -31.9518)
perth = coords |> filter(distances == min(distances))

pp = data |> filter(lon == perth$lon & lat == perth$lat)
perth_distance = perth$distances
perth_distance


distances = c(brisbane_distance, perth_distance, sydney_distance, melbourne_distance) |> sort()
distances

pp1 = pp[1, ]
Perth = data.frame(lon = pp1$lon, lat = pp1$lat, Individuals = 1, locations = "Perth")


coordinates(Perth) = ~ lon + lat
proj4string(Perth)=proj4string(AllData) # set it to lat-long
Perth = spTransform(Perth,proj4string(AllData))
e <- extent(110, 155, -45, -9)
r <- raster(e, ncols = 10000, nrows = 10000)
# you need to provide a function 'fun' for when there are multiple points per cell
Perth2 <- rasterize(Perth, r, fun = mean)
Perth3 = raster::resample(Perth2[[2]], raster(ext= e, resolution= 1,crs=projection(Perth)))
raster::writeRaster(Perth3, format="ascii", filename = "Inputs/intro_Perth_port.asc", NAflag = -9, overwrite = T, datatype = "UTF-8")
```

Visualization of the survival probabilities and the introduction points for cakile edentula.

```{r}
a3 = rbind(data.frame(lon = ss$lon, lat = ss$lat, Individuals = 1, locations = rep("Syd", 1)),
           data.frame(lon = mm$lon, lat = mm$lat, Individuals = 1, locations = rep("Mel", 1)),
           data.frame(lon = pp$lon, lat = pp$lat, Individuals = 1, locations = rep("Perth", 1)),
           data.frame(lon = bb$lon, lat = bb$lat, Individuals = 1, locations = rep("Bris",1)))

lp = raster("Inputs/climate_suitability2.asc") |> rasterToPoints() |> as.data.frame()  |> rename(value = climate_suitability2) 
ggplot( ) + 
  geom_tile(data = lp , aes(x = x, y = y, fill = value)) + 
  scale_fill_gradient("suitability", low = "white", high = "black") +
  theme_classic() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + 
  geom_point(data = a3, aes(x = lon, y = lat, color = locations))
```


4. Define the *landscape parameter* in RangeshiftR for simulation with the climate map and the species distribution and the parameters:

      - resolution = $1km$. 
      - HabPercent = TRUE: since the data in our suitability map is a probability value between 0 and 1. 
      - K_or_DensDep = 10000: to infer that a location with a probability of 1 would have a carrying capacity of 10000. The carrying capacities for all locations are calculated based on the probability value and the 100 threshold. 

5. We define the Species parameters which include:

- *demography*: Assumptions include:

    - the population dynamics are described by a female-only model 
    - generations are non-overlapping 
    - the maximum growth rate = 1000 : mean number of offspring per female and reproductive event at very low density. (AM Payne, 1984: "Reproduction ans survivorship of cakile edentula var.lacutris along the Lake Huron shoreline" suggests that "The largest number of of upper fruit segments on a mature plant was 1723 or a potential 3446 seeds")
    - bc = 0.000001 : competition coefficient (under-compensatory)

- *dispersal*: This contains information on *Emigration* probability, *Transport* modeled with a dispersal kernel and *Settlement* whose default options assume that an individual will die if it arrives in an unsuitable cell and settle if it’s suitable. The assumption:

    - Emigration probability = 0.1
    - *short range* and *long range transfer* at $20,000m$ with $probability = 0.5$ and $100,000m$ with $probability = 0.1$ respectively (You provided this values). This suggests that we have a 3rd possible dispersal distance equal to $0$ which has a probability of $1 - (0.5 + 0.1) = 0.4$. However, rangeshiftR  provides for either a single or double kernel to define distance. To fit this, I considered the odds between the short dispersal distance (20km) and the long distance (100km) and used this to recalculate their probabilities assuming the 3rd option of $0Km$ is not present.

$$P_{20km} = P_{20km}/{(P_{20km} + P_{100Km})} = 0.83$$

$$P_{100km} = P_{100km}/{(P_{20km} + P_{100Km})}=0.17$$

*Settlement* is defined with the parameters:

    - MaxSteps = 10: Maximum number of steps before mortality
    - Settle = 2: if cell is not suitable, individual settles in one of the suitable cells around the cell it landed (relaxed settlement condition)

6. Initialize the parameters for simulation

7. Simulation parameters

We will run $1\ experiments$ with $10\ Replicates$ simulating the spread of cakile edentula over $1000\ years$ and record the *population*, *occurance* and *range* every $1\ years$

8. Combine all the predefined parameters into a master object and run the simulation. We will run two separate simulations, one for the short range dispersal and the other for the long range dispersal.

9. Run simulations

# default simulation parameters

The values for each parameter are sampled randomly from the limits: carrying_capacity (1000 - 10000), max_offspring (10 - 1000), competition_coefficient  (0.00000005 - 0.00005), short_range_distance (5000m - 30000m), long_range_distance (70000m - 200000m), short_range_probability (0.5 - 0.9), emigration_probability (0.0 - 0.5), no_introduced_per_cell (10 - carrying capacity), Year (100 years - 1000 years) and 
Species (10 - 1000). 


```{r, default_parameters}
parameters <- function (index) {
  tibble::tibble(intro_location = intro_location[index],
                 intro_suitability = intro_suitability[index],
                 carrying_capacity = sample(1000:10000, 1), 
                  max_offspring = sample(10:1000, 1),
                  competition_coefficient = sample(seq(0,0.005, 0.00000001), 1),
                  short_range_distance = sample(1000:30000, 1),
                  long_range_distance = sample(30001:200000, 1),
                  short_range_probability = sample(seq(0.5,0.9, 0.01), 1),
                  emigration_probability = sample(seq(0.0,0.5, 0.01), 1),
                  no_introduced_per_cell = sample(10:carrying_capacity, 1), 
                  Year = sample(100:1000, 1),
                  Species = sample(10:1000, 1))
}
```

# Summary of the results

```{r}
# database table to store simulation output
library(DBI)
library(RPostgres)

#provide the name of your db
db <- "postgres" 
host_db <- "postgres.cl0erzvvnfux.ap-northeast-1.rds.amazonaws.com"   
db_port <- 5432  
db_user <- "philipp"  
db_password <- '****' # enter the password here

con <- dbConnect(RPostgres::Postgres(), dbname = db, host=host_db, port=db_port, user=db_user, password=db_password) 


## simulation
source("simulation_function.R")

intro_location = c("Brisbane", "Melbourne", "Perth", "sydney")

intro_suitability = c(bb$Unknown.species_AllData_Full.logistic.values, mm$Unknown.species_AllData_Full.logistic.values, pp$Unknown.species_AllData_Full.logistic.values, ss$Unknown.species_AllData_Full.logistic.values)

n = 111

for(i in sample(1:length(intro_location), n, replace = T)){
  my_simulation(parameters = parameters(i))
  dbGetQuery(conn = con, "SELECT COUNT(*) FROM simulated_population_lag_table_4_locations")
}
```









